CALVIS: chest, waist and pelvis circumference from 3D human body meshes as ground truth for deep learning

### 2019-10-28 PaperRMFI2019/samplepaper.tex
related work and cnn

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads, orivec]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{float}
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
\renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{CALVIS: chest, waist and pelvis circumference from 3D human body meshes	
as ground truth for deep learning}
%
\titlerunning{Chest, wAist and peLVIS circumference from 3D human body meshes}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Yansel Gonzalez Tejeda\orcidID{0000-0003-1002-3815} \and
Helmut Mayer\orcidID{1111-2222-3333-4444}}
%
\authorrunning{Gonzalez-Tejeda and Mayer}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Paris Lodron University of 
	Salzburg, Kapitelgasse 4-65020 Salzburg, Austria
\url{https://www.uni-salzburg.at}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
In this paper we present CALVIS, a method to calculate \textbf{C}hest, 
w\textbf{A}ist and pe\textbf{LVIS} circumference from 3D human body meshes. Our 
motivation is to use 
this data as ground truth for training convolutional neural networks 
(CNN). Previous work had used the large scale CAESAR dataset or determined 
these anthropometrical measurements $\textit{manually}$ from a person 
or human 3D body meshes. 
Unfortunately, acquiring these data is a cost and 
time consuming endeavor. In contrast, our method can be used on 
3D meshes automatically. We synthesize eight human body meshes and 
apply 
CALVIS to calculate chest, waist and pelvis circumference. We evaluate the 
results qualitatively and observe that the measurements can indeed be used 
to estimate the shape of a person. We then asses the 
plausibility of our approach by exploiting the calculated human dimensions as 
ground truth to train a small CNN. After
having trained the network with 
our data, we achieve competitive prediction error.
Furthermore, we make the implementation of CALVIS publicly available to 
advance the field.

\keywords{chest, waist and pelvis circumference \and 3D human body mesh \and 
deep learning}
\end{abstract}
%
%
%
\section{Introduction}\label{sec:intro}
\textbf{Motivation} 
Predicting 3D human body measurements from images is crucial in several 
scenarios 
like virtual try-on, animating, ergonomics, computational forensics and 
even health and mortality estimation. Researches had 
named this 
measurements body intuitive controls \cite{Allen.2003}, biometric 
measurements \cite{Sigal.2008}, body dimensions 
\cite{DBLP:conf/bmvc/ChenRC11}, (European standard EN 13402), semantical 
parameters \cite{Yang.2014}, 
traditional anthropometric 
measurements 
\cite{Wuhrer2011} or only ``shape" as in human shape estimation 
\cite{Guan.2013,Bogo:ECCV:2016,Loper.2015,Dibra.2016a,Pishchulin.2017}. In 
contrast, we assume a more 
anthropometric approach 
motivated by comprehensive compendiums like Panero and Zelnik, 1979 
\cite{panero1979human}.
Throughout this paper the term human body 
dimensions 
will be used to refer to the above measurements and we will focus on three of 
these dimensions, namely chest, waist and 
pelvis circumference.

@@

3D human body generative models offers such an alternative. We propose to 
synthesize 3D human body meshes using the SMPL \cite{Loper.2015} generative 
model. Once we have 
the 3D meshes we can compute chest, waist and pelvis circumference. The next 
step after obtaining the 
measurements is to use a camera model to render
images. Finally, in possession of this ground truth we can 
input this images to the learning algorithm and perform inference.

\textbf{Contributions} In summary, our contributions are $1)$ CALVIS: a method 
capable of automatically outputting chest, waist and pelvis circumference from 
3D body meshes. $2)$ A prototype deep learning framework (CNN) with synthetic 
images of 3D human body meshes as input and the human body dimension obtained 
with CALVIS as supervision signal.

\section{Related Work}\label{sec:rel_work}

\subsection{Human Body Data synthesis}
The \cite{Anguelov.2005} SCAPE model opened wide possibilities in the field  
because it provided the community with a generative model of the human 
body. However, if (Problem they address) (which method?)(Results?) (Open 
Problems).
After some(which, reference here) attempts on improving (say here either 
quality or spped), Loper et al., 2018 \cite{Loper.2015} developed a generative 
human body model.(Problem they address) (which method?)(Results?) (Open 
Problems).
Pischchulin et al., 2017 \cite{Pishchulin.2017} synthesize data. (Problem they 
address) (which method?)(Results?) (Open Problems).
More recently, Joo et al. \cite{Joo.2018} presents a human model with SMPL as a 
body model and added hand and face. We do not use that more complex model in 
this paper, that is future work.

\subsection{Human Body Dimensions from 3D models}
Allen et al., 2003 \cite{Allen.2003} proposed how to relate what they call body 
intuitive controls, e.g. height, weight, age and sex with PCAs performed on the 
space X (replace X!) using a linear mapping. Describe more please! (Problems)- 
make connection to next related work.  
Sigal et al., 2008 \cite{Sigal.2008} used a generative model (which method?) to 
estimate what he calls biometric measurements. (Results?) (Open Problems).
It is obvious that the problem of estimating the shape and pose of a human 
having only an image it is an under-constrained  (or inverse) problem. 
\cite{Guan.2013}  argues that (why) this situation can be alleviated by 
providing the shape model with the capability of keeping the height constant. 
This is indeed not the case for the original SCAPE model because the height is 
a semantical parameter and semantical parameters are “spread” over all bases. 
The arguments goes like this: SCAPE is a good human shape model. At its core 
this model relates triangles deformations to semantical parameters. One of the 
first problem one encounters when trying to model the human shape is the model 
size: because one needs a lot (~6000) triangles to sculpt a human body, you 
have the same amount of dimensions. Then one applies dimensionality reduction 
techniques, in this case PCA, having at the end around 20 components. These 
components do not describe intuitively (are not expressed explicit in the model 
) human shape characteristics (e.g., height, chest circumference or waist), 
nevertheless they are the bases of a vector space containing human shapes. Next 
\cite{Guan.2013} goes an argue that by “rotating” one (or more) of these bases 
we can vary the other bases  without being forced to vary the height.

Wuhrer and Shu, 2013 \cite{Wuhrer.2013} proposed a method to 
create/estimate/generate 3D human shapes from traditional anthropometric 
measurements. They learned a correlation between the 3D bodies meshes and 1D 
measurements. Then, when predicting a new shape, they find an initial solution 
based on the learned correlation (yansel WHICH correlation) and refine this 
solution by a two-step non linear optimization. They build up on two 
assumptions, namely, that the initial estimation should be in the plausible 
space of human shapes and second, that, in order to generalized, the model has 
to be refined without using this prior knowledge. While the first step use 
information from the learned shape prior, the second step introduce a 
smoothness constraint on the objective function. Like this work we use a set of 
traditional anthropometric measurements, unlike them we calculate 1D 
measurements from 3D human bodies to use them as ground truth for later 
inference. Their approach is able to capture variation outside the shape space 
spanned by the training data, in other words, the approach generalized well.
At the same time, Boisvert et al., 2013 \cite{Boisvert.2013} estimate human 
shape from measurements (what is that)(which method?)(Results?) (Open 
Problems). These two works can not be directly quantitatively compared because 
they use different set of measurements.
Moreover, Yang et al., 2014 \cite{Yang.2014} developed a method to map 
principal component to what they call semantical parameters. These semantical 
parameter are indeed anthropometric measurements, which justify their naming: 
in contrast to anthropometric meaningless principal components, their model is 
capable of manipulating concepts like chest and waist 
circumference.(Method?)(Results?).   
Tsoli et al., 2014 \cite{Tsoli.2014} extract what they calls anthropometric or 
tailored measurements introduced model-based anthropometry (what is that)(which 
method?)(Results?) (Open Problems).

@@

\subsection{Human Shape Estimation with Artificial Neural Networks}
A huge amount of research has been conducted in recent years to address the 
problem of 3D/2D human shape estimation using ANNs. Experimented with different 
architectures.
An state-of-the-art method is \cite{kanazawaHMR18} where they estimate human 
shape and pose.

@@

We 
can formalize this intuition by considering the cross-sectional length of the 
2D intersection curves\cite{yansel} along the y-axis. Moreover, we 
can 
use a plane $\boldsymbol{\pi_j}$ parallel (with normal $\vec{n}_{||}$) to the 
floor to intersect the mesh at point $q_j = (0, q^y_j, 0), \,  q^y_j \in 
\mathbb{R}$ along the y-axis. Since $\mathcal{M}$ is triangulated, the boundary 
of this 
intersection is a polygonal curve consisting in segments with length $s_i$. 
Therefore, we 
can 
determine the intersection boundary length as
\begin{equation}\label{eq:boundary_length}
\mathcal{BL}(\mathcal{M}, \vec{n}_{||}, q) = \sum_{i = 
	1}^{i = N}s_i
\end{equation}

@@

\begin{figure}[t]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{Figure_1.eps}
	\end{center}
	\caption{Human Body Mesh Signature for male and female meshes. The 
		function resembles a rotated silhouette of the human body and exhibits 
		several \textit{extrema}. See section 
		\ref{subsec:mesh_signature_extrema} for discussion of these extrema.}
	\label{fig:hbm_signature}
\end{figure}

@@

We conduct two experiments. In the first experiment we synthesize eight (four 
female and four male) human body meshes using shape parameters provided by 
SURREAL \cite{varol17_surreal}.
The meshes reflect human figures characteristics such as bulky, slim, small and 
tall subjects. Then we apply CALVIS to calculate chest, waist and pelvis 
circumference. Since we do not have ground truth, we evaluate the results 
qualitatively by comparing our method 
with \cite{Dibra.2016b}.
The second 
experiment serves to asses the plausibility of our approach to use the 
synthetic data for deep learning. We input the calculated human 
dimensions to an artificial neural network. After having trained the network 
with our data, we show that learning converges.

@@

\subsection{Learning with Calvis-Net}\label{subsec:learning}
This experiment demonstrates the plausibility of employing calvis to obtain 
data 
that can be used as a ground truth for machine learning (in general) and 
specifically as supervising signal for CNN.

\subsubsection{Experimental setup} We annotate the CMU dataset with calvis. The 
automatic annotation process took approximately $1\,h\, 47\, min$ in an 
enhanced 
modern personal computer.

@@

%------------------------------------------------------------------------
\section{Conclusion}

The method operates optimally.
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{egbib}
%

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads, orivec]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{float}
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\usepackage[pagebackref=true,breaklinks=true,colorlinks,bookmarks=false]{hyperref}
\renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{CALVIS: chest, waist and pelvis circumference from 3D human body meshes	
as ground truth for deep learning}
%
\titlerunning{Chest, wAist and peLVIS circumference from 3D human body meshes}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Yansel Gonzalez Tejeda\orcidID{0000-0003-1002-3815} \and
Helmut Mayer\orcidID{1111-2222-3333-4444}}
%
\authorrunning{Gonzalez-Tejeda and Mayer}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Paris Lodron University of 
	Salzburg, Kapitelgasse 4-65020 Salzburg, Austria
\url{https://www.uni-salzburg.at}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
In this paper we present CALVIS, a method to calculate \textbf{C}hest, 
w\textbf{A}ist and pe\textbf{LVIS} circumference from 3D human body meshes. Our 
motivation is to use 
this data as ground truth for training convolutional neural networks 
(CNN). Previous work had used the large scale CAESAR dataset or determined 
these anthropometrical measurements $\textit{manually}$ from a person 
or human 3D body meshes. 
Unfortunately, acquiring these data is a cost and 
time consuming endeavor. In contrast, our method can be used on 
3D meshes automatically. We synthesize eight human body meshes and 
apply 
CALVIS to calculate chest, waist and pelvis circumference. We evaluate the 
results qualitatively and observe that the measurements can indeed be used 
to estimate the shape of a person. We then asses the 
plausibility of our approach by generating
ground truth with CALVIS to train a small CNN. After
having trained the network with 
our data, we achieve competitive validation error.
Furthermore, we make the implementation of CALVIS publicly available to 
advance the field.

\keywords{chest, waist and pelvis circumference \and 3D human body mesh \and 
deep learning}
\end{abstract}
%
%
%
\section{Introduction}\label{sec:intro}
\textbf{Motivation} 
Predicting 3D human body measurements from images is crucial in several 
scenarios 
like virtual try-on, animating, ergonomics, computational forensics and 
even health and mortality estimation. Researches had 
named these
measurements body intuitive controls \cite{Allen.2003}, biometric 
measurements \cite{Sigal.2008}, body dimensions 
\cite{DBLP:conf/bmvc/ChenRC11}, (European standard EN 13402), semantical 
parameters \cite{Yang.2014}, 
traditional anthropometric 
measurements 
\cite{Wuhrer2011} or only ``shape" as in human shape estimation 
\cite{Guan.2013,Bogo:ECCV:2016,Loper.2015,Dibra.2016a,Pishchulin.2017}. In 
contrast, we assume a more 
anthropometric approach 
motivated by comprehensive compendiums like Panero and Zelnik, 1979 
\cite{panero1979human}.
Throughout this paper the term human body 
dimensions 
will be used to refer to the above measurements and we will focus on three of 
these dimensions, namely chest, waist and 
pelvis circumference.

@@

3D human body generative models offers such an alternative. In the line of previous work, we start by synthesizing 3D human body meshes using the SMPL \cite{Loper.2015} generative 
model. Unlike other works, once we have 
the 3D meshes we compute chest, waist and pelvis circumference. The next 
step after obtaining the 
measurements is to use a camera model to render
images. Finally, in possession of this ground truth we can 
input these images to the learning algorithm and train with the human body dimensions as supervision signal.

\textbf{Contributions} In summary, our contributions are $1)$ CALVIS: a method 
capable of automatically outputting chest, waist and pelvis circumference from 
3D body meshes. $2)$ A prototype deep learning framework (CNN) with synthetic 
images of 3D human body meshes as input and the human body dimensions obtained 
with CALVIS as supervision signal.

\section{Related Work} \label{sec:related_work}

%-------------------------------------------------------------------------
\subsection{Human Dimensions Estimation from Images.}
Using metrology techniques \cite{ams.2008.BenAbdelkaderY08} and mixture of experts model \cite{Sigal.2008} human body dimensions like height, arm span and even weight had been estimated from images. Chest and waist size had been considered as well \cite{Guan.2013}. These investigations used human dimensions for validation (ground truth) that were directly recovered from individuals. In contrary, we calculate them  
consistently from 3D meshes.
Another significant research direction studies shape as (non necessarily meaningful) parameters of some model. For example \cite{Bogo:ECCV:2016} claim describing the first method 
to 
estimate pose and shape from a single image, while \cite{DBLP:Lassner2017} estimates shape from 91 keypoints using decision 
forests. Shape is understood as the 
parameters of the SMPL model but not human dimensions.

@@

\subsection{Human Body Data Synthesis.}
The SCAPE model\cite{Anguelov.2005} opened wide possibilities in the field. It 
provided the scientific community with a method to build realistic 3D human 
body meshes of different shapes in different poses. In order to synthesize a 
human body mesh a template body must be deformed. The model exhibits, however, an important limitation, namely the last 
step on the pipeline is an expensive optimization problem. Other researchers had used variation of the SCAPE model (S-SCAPE \cite{Pishchulin.2017}) to synthesize human bodies but focused on people detection.

After some attempts on improving the human models quality, for example, to make 
more 
easily capturing the correlation between body shape and pose 
\cite{HaslerSSRS09}, or to better understand body dimensions (SPRING model 
\cite{Yang.2014}), Loper et al., 2015 \cite{Loper.2015} developed the SMPL generative 
human body model. In this work we synthesize 3D human meshes using this 
model. We will briefly describe the model in subsection \ref{subsec:smpl_model}.

Our approach to synthesize images from 3D meshes has been strongly influence by the recent publication of the large scale dataset SURREAL \cite{varol17_surreal}. This work uses the SMPL model 
to generate images 
from humans with random backgrounds. However, no human body dimensions are computed of estimated.

More recently, a human model with added hands and face had been presented \cite{Joo.2018}. We do not use this more complex model 
because the human dimensions we estimate do not require that level 
of detail.

\subsection{Human Body Dimensions from 3D Models.}
Although extensive research has been carried out on human shape estimation, 
methods to consistently define shape based on 3D mesh 
anthropometric measurements have been little explored. Only a handful of researchers 
have 
reported calculating 1D body dimensions from 3D triangular meshes to use them as ground truth for training and validation in a variety of inference processes.

Early research performed feature analysis on what they 
call body intuitive controls, e.g. height, weight, age and gender \cite{Allen.2003}  but they do not 
calculate them. Instead they use the CAESAR demographic data. Recording human dimensions beyond height like body fat and the more abstract ``muscles" are described by \cite{HaslerSSRS09}.

Pertinent to our investigation is also the inverse problem: generating 3D human 
shapes from traditional anthropometric measurements \cite{Wuhrer.2013}. Like this work we use a set of 
anthropometric measurements that we call dimensions, unlike them we calculate 
1D measurements from 3D human bodies to use them as ground truth for later 
inference.

Strongly related to our work are methods that calculate waist and 
chest circumference by slicing the mesh on a fixed plane and compute the convex hull of the contour \cite{Guan.2013} or path length from identified (marked) 
vertices on the 3D mesh \cite{Boisvert.2013,Dibra.2016a,Dibra.2016b}. However, is not clear how they define the human body dimensions.

In contrast, we do not 
calculate the dimensions from fix vertices on the template 
mesh. Instead, we 
take a more anthropometric approach and use domain knowledge to calculate these 
measurements.

\subsection{Human Shape Estimation with Artificial Neural Networks}
A huge amount of research has been conducted in recent years to address the 
problem of 3D/2D human shape estimation using ANNs. A state-of-the-art method 
is \cite{kanazawaHMR18} where they estimate human 
shape and pose. While these ANN's output are human body models parameters 
(i.e., $\beta$s in \cite{Dibra.2016a} and \cite{varol18_bodynet}) our network 
is 
capable to output human dimensions directly.

@@

We 
can formalize this intuition by considering the cross-sectional length of the 
2D intersection curves along the y-axis. Moreover, we 
can 
use a plane $\boldsymbol{\pi_j}$ parallel (with normal $\vec{n}_{||}$) to the 
floor to intersect the mesh at point $q_j = (0, q^y_j, 0), \,  q^y_j \in 
\mathbb{R}$ along the y-axis. Since $\mathcal{M}$ is triangulated, the boundary 
of this 
intersection is a polygonal curve consisting in segments with length $s_i$. 
Therefore, we 
can 
determine the intersection boundary length as
\begin{equation}\label{eq:boundary_length}
\mathcal{BL}(\mathcal{M}, \vec{n}_{||}, q) = \sum_{i = 
	1}^{i = N}s_i
\end{equation}

@@

\begin{figure}[t]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{Figure_1.eps}
	\end{center}
	\caption{Human Body Mesh Signature for male and female meshes. The 
		function resembles a rotated silhouette of the human body and exhibits 
		several \textit{extrema}.}
	\label{fig:hbm_signature}
\end{figure}

@@

We conduct two experiments. In the first experiment we synthesize eight (four 
female and four male) human body meshes using shape parameters provided by 
SURREAL \cite{varol17_surreal}.
The meshes reflect human figures characteristics such as bulky, slim, small and 
tall subjects. Then we apply CALVIS to calculate chest, waist and pelvis 
circumference. Since we do not have ground truth, we evaluate the results 
qualitatively.
The second 
experiment serves to asses the plausibility of our approach to use the 
synthetic data for deep learning. We input the calculated human 
dimensions to an artificial neural network. After having trained the network 
with our data, we show that the network is able to learn the task.

@@

\subsection{Learning with Calvis-Net}\label{subsec:learning}
This experiment demonstrates the plausibility of employing CALVIS to obtain 
data 
that can be used as a ground truth for machine learning (in general) and 
specifically as supervising signal for CNN.

\subsubsection{Experimental setup} We annotate the CMU dataset with CALVIS. The 
automatic annotation process took approximately $1\,h\, 47\, min$ in an 
enhanced 
modern personal computer.

@@

%------------------------------------------------------------------------
\section{Conclusion}
In this paper we presented CALVIS, a method to calculate chest, waist and pelvis circumference from 3D human meshes. 
We demonstrate that our approach can be used to train a CNN by imputing synthetic images of humans and using the measurements calculated with CALVIS as supervision signal.
Furthermore, we contribute with a prototype CNN CALVIS-NET. Our experiments show that the CNN is able to learn how to perform this task, achieving mean absolute validation error of 0.061698 m plus +-.
The code 
and data use 
in this paper will be made publicly available for researchers.
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{egbib}
%
